Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-01-24 paper-new/categorical-model.tex
updates to abstract categorical model section

In summary, an extensional model consists of:
\begin{enumerate}
  \item A CBPV model internal to reflexive graphs.
  \item Composition and identity on value and computation relations that form a category.
  \item An identity-on-objects functor $\upf : \mathcal V_e \to \mathcal V_f$ taking each value relation to a morphism that left-represents it.
  \item An identity-on-objects functor $\dnf : \mathcal E_e^{op} \to \mathcal E_f$ taking each computation relation to a morphism that right-represents it.
  \item The CBPV connectives $U,F,\times,\to$ are all \emph{covariant} functorial on relations
    \begin{itemize}
    \item $U(dd') = U(d)U(d')$
    \item $F(cc'') = F(c)F(c')$
    \item $(cc') \to (dd') = (c \to d)(c' \to d')$
    \item $(c_1c_1') \times (c_2c_2') = (c_1 \times c_2)(c_1'\times c_2')$
    \end{itemize}
  \item A distinguished value type $D$ with distinguished relation
    $\injarr{}: U(D \to F D) \rel D$ satisfying $\dnc {\injarr{}}F(\upc{\injarr{}}) = \id$
\end{enumerate}

@@

A second consequence of working intensionally is that the squares in the representable
properties must now involve a notion of ``delay" or ``perturbation'' in order to
keep the function morphisms on each side in lock-step. Intuitively, the perturbations
have no effect other than to cause the function to which they are applied to ``wait''
in a specific manner.
We formalize this notion by requiring that for each object $A$ in $\vf$,
there is a monoid of \emph{perturbations} $P^V_A$ and a monoid homomorphism
$\ptbv_A : \pv_A \to \vf(A,A)$.
Similarly, for each $B : \ef$ there is a monoid $P^C_B$ and a
homomorphism $\ptbe_B : \pe_B \to \ef(B,B)$.
If $\delta \in P^V_A$, we will sometimes omit the homomorphism $\ptbv_A$ and simply write
$\delta$ to refer to the morphism $\ptbv_A(\delta) \in \vf(A,A)$, and likewise
for computation perturbations. The context will make clear whether we are referring
to an element of the perturbation monoid or the corresponding morphism.

@@

The perturbations must be preserved by $\times$, $\arr$, $U$, and $F$.

@@

Lastly, we require that all perturbations be weakly bisimilar to the identity morphism,
capturing the notion that they have no effect other than to delay. We observe that
the set of endomorphisms $f$ such that $f$ is weakly bisimilar to the identity
forms a monoid under composition.

@@

\begin{definition}\label{def:step-2-model}
  A \emph{step-2} model of intensional gradual typing consists of all the data of a step-1 model plus:
  \begin{enumerate}
    \item $\pv_A$ and a monoid homomorphism 
      \[ \ptbv_A : \pv_A \to \{ f \in \vf(A,A) \mid f \bisim \id \} \]
    \item $\pe_B$ and a monoid homomorphism 
      \[ \ptbe_B : \pe_B \to \{ g \in \ef(B,B) \mid g \bisim \id \} \]
    \item The functors $\times$, $\arr$, $U$, and $F$ preserve perturbations.
    \item The push-pull property holds for all $c : A \rel A'$ and all $d : B \rel B'$.
  \end{enumerate}
\end{definition}

@@

\begin{definition}\label{def:quasi-left-representable}
Let $c : A \rel A'$ be a value relation. We say that $c$ is \emph{quasi-left-representable by}
$f \in \vf(A, A')$ if there are perturbations $\delta_c^{l,e} \in \pv_A$ and
$\delta_c^{r,e} \in \pv_{A'}$ such that the following squares commute:

\begin{center}
  \begin{tabular}{ m{7em} m{7em} } 
    % UpL
    \begin{tikzcd}[ampersand replacement=\&]
      A \& {A'} \\
      {A'} \& {A'}
      \arrow["f"', from=1-1, to=2-1]
      \arrow["\delta_c^{r,e}", from=1-2, to=2-2]
      \arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-2]
      \arrow[from=2-1, to=2-2, Rightarrow, no head]
    \end{tikzcd}
    &
    % UpR
    \begin{tikzcd}[ampersand replacement=\&]
      A \& {A} \\
      {A} \& {A'}
      \arrow["\delta_c^{l,e}"', from=1-1, to=2-1]
      \arrow["f", from=1-2, to=2-2]
      \arrow[from=1-1, to=1-2, Rightarrow, no head]
      \arrow["c"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
    \end{tikzcd}
  \end{tabular}
\end{center}
\end{definition}

@@

\begin{center}
  \begin{tabular}{ m{7em} m{7em} } 
    % DnR
    \begin{tikzcd}[ampersand replacement=\&]
      {B} \& {B'} \\
      {B} \& {B}
      \arrow["\delta_d^{l,p}"', from=1-1, to=2-1]
      \arrow["g", from=1-2, to=2-2]
      \arrow["R", "\shortmid"{marking}, no head, from=1-1, to=1-2]
      \arrow[from=2-1, to=2-2, Rightarrow, no head]
    \end{tikzcd}
    &
    % DnL
    \begin{tikzcd}[ampersand replacement=\&]
      {B'} \& {B'} \\
      {B} \& {B'}
      \arrow["g"', from=1-1, to=2-1]
      \arrow["\delta_d^{r,p}", from=1-2, to=2-2]
      \arrow[from=1-1, to=1-2, Rightarrow, no head]
      \arrow["R"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
    \end{tikzcd}
  \end{tabular}
\end{center}
\end{definition}

Besides the perturbations, one other difference between the extensional
and intensional versions of the representability axioms is that in the
extensional setting, the rules build in the notion of composition, whereas
their intensional counterparts do not.
In the extensional setting, we do not have horizontal composition of squares, which
is required to derive the versions of the rules that build in composition
from the versions that do not.
In the intensional setting, we do have horizontal composition of squares,
so we can take the simpler versions as primitive and derive the ones
involving composition (though we must be careful about the perturbations
when doing so!).

\begin{definition}\label{def:step-3-model}
  A \emph{step-3 intensional model} consists
  of all the data of a step-2 intensional model, with the following additional data:
  \begin{enumerate}
    \item Functors $\upf : \ve \to \vf$ and $\dnf : \ee^{op} \to \ef$ % TODO: image is thin?
    \item Every value edge $c : A \rel A'$ is quasi-left-representable by $\upf(c)$ and
    every computation edge $d : B \rel B'$ is quasi-right-representable by $\dnf(d)$.
  \end{enumerate}
\end{definition}

@@

Now we can discuss what it means for an intensional model to model the dynamic type.
This applies to any of the above abstract model definitions, i.e., steps 0-3.

\begin{definition}\label{def:step-4-model}
  A \emph{step-$i$ intensional model with dyn} is a step-$i$ model $\mathcal M$ such that:
  %a distinguished value object $D \in \ob(\vf)$ such that:
  %
  \begin{enumerate}
    \item There is a distinguished value object $D \in \ob(\vf)$.

@@

    \item If $c : A \rel A'$, then $\text{inj}_{A} = c \comp \text{inj}_{A'}$.
  \end{enumerate}
\end{definition}

@@

Suppose we have a \hyperref[def:step-1-model]{step-1 intensional model} $\mathcal{M}$ with dyn.
Recall that a step-1 intensional model consists of a step-0 model (i.e., a
category internal to the category of CBPV models), along with the necessary
categories and functors for bisimilarity as discussed in Section
\ref{sec:abstract-model-bisimilarity}.
Further, recall that a \hyperref[def:step-2-model]{step-2 model} has everything
a step-1 model has, with the addition of perturbation monoids $\pv_A$ for all
$A$ and $\pe_B$ for all $B$.
These perturbations must be interpretable via homomorphisms
$\ptb_A$ and $\ptb_B$ as endomorphisms bisimilar to the identity,
and moreover, the push-pull property must hold for all
value relations $c$ and all computation relations $d$.

% TODO do we really need all of this data?
What ``external" data do we need to define such a model?
Suppose we are given, for each value type $A$, a monoid $\pv_A$ with homomorphism
$\ptb_A$ into the endomorphisms on $A$ bisimilar to the identity, and likewise, for
each computation type $B$, a monoid $\pv_B$ and homomorphism $\ptb_B$.
Furthermore, suppose we are given, for each value type $A$, a distinguished endomorphism
$\delta_A \in \ef(FA, FA)$ such that $\delta_A \bisim \id_{FA}$.
Finally, suppose that each value relation $c : A \rel A'$ satisfies the push-pull property
and likewise for each computation relation $d : B \rel B'$.

Then we claim that from the above data, we can construct a step-2 intensional model.
For the proof, see Lemma \ref{lem:step-1-model-to-step-2-model} in the Appendix.

@@

Now suppose we have a step-2 intensional model $\mathcal{M}$.
We will describe how to construct a \hyperref[def:step-3-model]{step-3 intensional model} $\mathcal{M'}$.

We now describe the construction of a step-3 model $\mathcal M'$.

@@

\subsubsection{Defining an Extensional Model}

Finally, suppose $\mathcal M$ is a step-3 intensional model.
We now describe how to build an extensional model.

@@

Next, we check that the requirements of an extensional model are satisfied.
In particular, we need to verify the representability properties.
We define functors $\upf$ and $\dnf$

>>>>

In summary, an extensional model consists of:
\begin{enumerate}
  \item A CBPV model internal to reflexive graphs.
  \item Composition and identity on value and computation relations that form a category.
  \item An identity-on-objects functor $\upf : \mathcal V_e \to \mathcal V_f$ taking each value relation to a morphism that left-represents it.
  \item An identity-on-objects functor $\dnf : \mathcal E_e^{op} \to \mathcal E_f$ taking each computation relation to a morphism that right-represents it.
  \item The CBPV connectives $U,F,\times,\to$ are all \emph{covariant} functorial on relations
    \begin{itemize}
    \item $U(dd') = U(d)U(d')$
    \item $F(cc') = F(c)F(c')$
    \item $(cc') \to (dd') = (c \to d)(c' \to d')$
    \item $(c_1c_1') \times (c_2c_2') = (c_1 \times c_2)(c_1'\times c_2')$
    \end{itemize}
  \item A distinguished value type $D$ with distinguished relation
    $\injarr{}: U(D \to F D) \rel D$ satisfying $\dnc {\injarr{}}F(\upc{\injarr{}}) = \id$
\end{enumerate}

@@

A second consequence of working intensionally is that the squares in the representable
properties must now involve a notion of ``delay" or ``perturbation'' in order to
keep the function morphisms on each side in lock-step. Intuitively, the perturbations
have no effect other than to cause the function to which they are applied to ``wait''
in a specific manner.
We formalize this notion by requiring that for each object $A$ in $\vf$,
there is a monoid of \emph{value perturbations} $\pv_A$ that is a submonoid of
$\{ f \in \vf(A,A) \mid f \bisim \id \}$.
Similarly, for each $B : \ef$ there is a monoid $\pe_B$ of
\emph{computation perturbations} that is a submonoid of 
$\{ g \in \ef(B,B) \mid g \bisim \id \}$.

% If $\delta \in P^V_A$, we will sometimes omit the homomorphism $\ptbv_A$ and simply write
% $\delta$ to refer to the morphism $\ptbv_A(\delta) \in \vf(A,A)$, and likewise
% for computation perturbations. The context will make clear whether we are referring
% to an element of the perturbation monoid or the corresponding morphism.

Note that we require that all perturbations be weakly bisimilar to the identity morphism,
capturing the notion that they have no effect other than to delay.
% We observe that
% the set of endomorphisms $f$ such that $f$ is weakly bisimilar to the identity
% forms a monoid under composition.

@@

The perturbations must be preserved by $\times$ $\timesk$, $\arr$, $\tok$, $U$, $\Uk$, $F$, and $\Fk$.

@@

@@

\begin{definition}\label{def:step-2-model}
  A \emph{step-2} model of intensional gradual typing consists of all the data of a step-1 model plus:
  \begin{enumerate}
    \item For each value type $A$, there is a monoid $\pv_A$ that is a submonoid of
    $\{ f \in \vf(A,A) \mid f \bisim \id \}$.
    \item For each computation type $B$, there is a monoid $\pe_B$ that is a submonoid of
    $\{ g \in \ef(B,B) \mid g \bisim \id \}$.
    \item For all $A$, the distinguished endomorphism $\delta_A^*$ is in $\pe_{FA}$.
    % \item $\pv_A$ and a monoid homomorphism 
    %   \[ \ptbv_A : \pv_A \to \{ f \in \vf(A,A) \mid f \bisim \id \} \]
    % \item $\pe_B$ and a monoid homomorphism
    %   \[ \ptbe_B : \pe_B \to \{ g \in \ef(B,B) \mid g \bisim \id \} \]
    \item The functors $\times, \timesk$, $\arr, \tok$, $U, \Uk$, $F, \Fk$ preserve perturbations.
    \item The push-pull property holds for all $c : A \rel A'$ and all $d : B \rel B'$.
  \end{enumerate}
\end{definition}

@@

% TODO: Give these squares names
\begin{definition}\label{def:quasi-left-representable}
Let $c : A \rel A'$ be a value relation. We say that $c$ is \emph{quasi-left-representable by}
$f \in \vf(A, A')$ if there are perturbations $\delta_c^{l,e} \in \pv_A$ and
$\delta_c^{r,e} \in \pv_{A'}$ such that the following squares commute:

\begin{center}
  \begin{tabular}{ m{7em} m{7em} } 
    % UpL
    \begin{tikzcd}[ampersand replacement=\&]
      A \& {A'} \\
      {A'} \& {A'}
      \arrow["f"', from=1-1, to=2-1]
      \arrow["\delta_c^{r,e}", from=1-2, to=2-2]
      \arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-2]
      \arrow[from=2-1, to=2-2, Rightarrow, no head]
    \end{tikzcd}
    &
    % UpR
    \begin{tikzcd}[ampersand replacement=\&]
      A \& {A} \\
      {A} \& {A'}
      \arrow["\delta_c^{l,e}"', from=1-1, to=2-1]
      \arrow["f", from=1-2, to=2-2]
      \arrow[from=1-1, to=1-2, Rightarrow, no head]
      \arrow["c"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
    \end{tikzcd}
  \end{tabular}
\end{center}

We call the first square $\upl$ and the second square $\upr$.

\end{definition}

@@

\begin{center}
  \begin{tabular}{ m{7em} m{7em} } 
    % DnR
    \begin{tikzcd}[ampersand replacement=\&]
      {B} \& {B'} \\
      {B} \& {B}
      \arrow["\delta_d^{l,p}"', from=1-1, to=2-1]
      \arrow["g", from=1-2, to=2-2]
      \arrow["R", "\shortmid"{marking}, no head, from=1-1, to=1-2]
      \arrow[from=2-1, to=2-2, Rightarrow, no head]
    \end{tikzcd}
    &
    % DnL
    \begin{tikzcd}[ampersand replacement=\&]
      {B'} \& {B'} \\
      {B} \& {B'}
      \arrow["g"', from=1-1, to=2-1]
      \arrow["\delta_d^{r,p}", from=1-2, to=2-2]
      \arrow[from=1-1, to=1-2, Rightarrow, no head]
      \arrow["R"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
    \end{tikzcd}
  \end{tabular}
\end{center}

We call the first square $\dnr$ and the second square $\dnl$.

\end{definition}

Besides the perturbations, one other difference between the extensional
and intensional versions of the representability axioms is that in the
extensional setting, the rules build in the notion of composition, whereas
their intensional counterparts do not.
In the extensional setting, we do not have horizontal composition of squares, which
is required to derive the versions of the rules that build in composition
from the versions that do not.
In the intensional setting, we do have horizontal composition of squares,
so we can take the simpler versions as primitive and derive the ones
involving composition.

Lastly, we require that the model satisfy a weak version of functoriality for 
the CBPV connectives $U,F,\times,\to$. 
First, we will need a definition:

\begin{definition}
  Let $c, c' : A \rel A'$. We say that $c$ and $c'$ are \emph{quasi-order-equivalent},
  written $c \qordeq c'$, if there exist perturbations $\delta^l_1, \delta^l_2 \in \pv_A$ and 
  $\delta^r_1, \delta^r_2 \in \pv_{A'}$ such that the following two squares exist:

  \begin{center}
    \begin{tabular}{ m{9em} m{9em} } 
      \begin{tikzcd}[ampersand replacement=\&]
        A \& {A'} \\
        A \& {A'}
        \arrow["\delta^l_1"', from=1-1, to=2-1]
        \arrow["\delta^r_1", from=1-2, to=2-2]
        \arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-2]
        \arrow["c'"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
      \end{tikzcd}
      &
      \begin{tikzcd}[ampersand replacement=\&]
        A \& {A'} \\
        A \& {A'}
        \arrow["\delta^l_2"', from=1-1, to=2-1]
        \arrow["\delta^r_2", from=1-2, to=2-2]
        \arrow["c'", "\shortmid"{marking}, no head, from=1-1, to=1-2]
        \arrow["c"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
      \end{tikzcd}
    \end{tabular}
  \end{center}

  We make the analogous definition for computation relations $d, d' : B \rel B'$.
\end{definition}

We require that the CBPV connectives $U,F,\times,\to$ are \emph{quasi-functorial} on relations,
which we specify as follows:
\begin{itemize}
  \item $U(d \comp d') \qordeq U(d)U(d')$
  \item $F(c \comp c'') \qordeq F(c)F(c')$
  \item $(cc') \to (dd') \qordeq (c \to d)(c' \to d')$
  \item $(c_1c_1') \times (c_2c_2') \qordeq (c_1 \times c_2)(c_1'\times c_2')$
\end{itemize}

We summarize the requirements of a step-3 model below:

\begin{definition}\label{def:step-3-model}
  A \emph{step-3 intensional model} consists
  of all the data of a step-2 intensional model, such that additionally:
  \begin{enumerate}
    \item There are functors $\upf : \ve \to \vf$ and $\dnf : \ee^{op} \to \ef$ % TODO: image is thin?
    \item Every value edge $c : A \rel A'$ is quasi-left-representable by $\upf(c)$ and
    every computation edge $d : B \rel B'$ is quasi-right-representable by $\dnf(d)$.
    \item The CBPV connectives $U,F,\times,\to$ are quasi-functorial on relations.
  \end{enumerate}
\end{definition}

% Want: U d \comp U d' = U(d \comp d')
%       F c \comp F c' = F(c \comp c')
% Add requirement: Either the model is functorial with respect to up/downcasts or with repsect to relations
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Now we can discuss what it means for an intensional model to model the dynamic type.
% This applies to any of the above abstract model definitions, i.e., steps 0-3.

\begin{definition}\label{def:step-4-model}
  % A \emph{step-$i$ intensional model with dyn} is a step-$i$ model $\mathcal M$ such that:
  A step-4 intensional model is a step-3 intensional model $\mathcal M$ such that:
  %a distinguished value object $D \in \ob(\vf)$ such that:
  %
  \begin{enumerate}
    \item There is a distinguished value object $D \in \ob(\vf)$.

@@

    %\item \eric{Do we need this?} If $c : A \rel A'$, then $\text{inj}_{A} = c \comp \text{inj}_{A'}$.
  \end{enumerate}
\end{definition}

@@

Suppose we have a \hyperref[def:step-1-model]{step-1 intensional model} $\mathcal{M}$.
Recall that a step-1 intensional model consists of a step-0 model (i.e., a
category internal to the category of CBPV models), along with the necessary
categories and functors for bisimilarity as discussed in Section
\ref{sec:abstract-model-bisimilarity}.
Further, recall that a \hyperref[def:step-2-model]{step-2 model} has everything
a step-1 model has, with the addition of perturbation monoids $\pv_A$ for all
$A$ and $\pe_B$ for all $B$.
Moreover, the push-pull property must hold for all
value relations $c$ and all computation relations $d$.

We claim that from a step-1 model, we can construct a step-2 model. 
% TODO give overview
For the proof, see Lemma \ref{lem:step-1-model-to-step-2-model} in the Appendix.

@@

Now suppose we have a step-2 intensional model $\mathcal{M}$.
We claim that we can construct a \hyperref[def:step-3-model]{step-3 intensional model} $\mathcal{M'}$.

% TODO give overview

For the proof, see Lemma \ref{lem:step-2-model-to-step-3-model} in the Appendix.

@@

\subsubsection{Defining an Extensional Model}\label{sec:extensional-model-definition}

Finally, suppose $\mathcal M$ is a \hyperref[def:step-4-model]{step-4 intensional model}
(i.e., a step-3 model with an interpretation of the dynamic type).
We now describe how to build an extensional model.

@@

The proof that this indeed defines an extensional model is given in the Appendix 
(see Section \ref{sec:extensional-construction-appendix}).

% Next, we check that the requirements of an extensional model are satisfied.
% In particular, we need to verify the representability properties.
% We define functors $\upf$ and $\dnf$

### 2024-01-24 paper-new/paper.tex
begin rearranging: discuss term syn/sem and then ordering syn/sem
begin description of denotational model for intensional gradual typing
work on abstract categorical model
lics format
lics review format
begin section on concrete model
begin discussion section
begin appendix
some intro work and add macros for comments
correct title/authors
updates to abstract categorical model section

\documentclass[acmsmall,screen]{acmart}
\usepackage{mathpartir}
\usepackage{tikz-cd}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{fancyvrb}
\usepackage{comment}

\usepackage{ stmaryrd }

@@

\title{Mechanized Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory}
\author{Eric Giovannini}
\author{Max S. New}

\begin{abstract}
  Gradually typed programming languages, which allow for soundly
  mixing static and dynamically typed programming styles, present a
  strong challenge for metatheorists. Even the simplest sound
  gradually typed languages feature at least recursion and errors,
  with realistic languages featuring furthermore runtime allocation of
  memory locations and dynamic type tags. Further, the desired
  metatheoretic properties of gradually typed languages have become
  increasingly sophisticated: validity of typed based equational
  reasoning as well as the relational property known as the gradual
  guarantee or graduality. Many recent works have tackled verifying
  these properties, but the resulting mathematical developments are
  highly repetitive and tedious, with few reusable theorems persisting
  across different developments.

@@

@@

\input{syntax}

\input{semantics}

@@

\section{Discussion}\label{sec:discussion}

>>>>

\documentclass[sigconf,anonymous,review,screen,9pt]{acmart}
\let\Bbbk\relax

\usepackage{quiver}
\usepackage{mathpartir}
% \usepackage{tikz-cd}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{fancyvrb}
\usepackage{comment}
\usepackage{array}

@@

\title{Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory}
\author{Eric Giovannini}
\affiliation{
  \department{Electrical Engineering and Computer Science}
  \institution{University of Michigan}
  \country{USA}
}
\email{ericgio@umich.edu}

\author{Tingting Ding}
\affiliation{
  \department{Electrical Engineering and Computer Science}
  \institution{University of Michigan}
  \country{USA}
}
\email{tingtind@umich.edu}

\author{Max S. New}
\affiliation{
  \department{Electrical Engineering and Computer Science}
  \institution{University of Michigan}
  \country{USA}
}
\email{maxsnew@umich.edu}

\begin{abstract}
  Gradually typed programming languages, which allow for soundly
  mixing static and dynamically typed programming styles, present a
  strong challenge for metatheorists. Even the simplest sound
  gradually typed languages feature at least recursion and errors,
  with realistic languages featuring furthermore runtime allocation of
  memory locations and dynamic type tags. Further, the desired
  metatheoretic properties of gradually typed languages have become
  increasingly sophisticated: validity of type based equational
  reasoning as well as the relational property known as the gradual
  guarantee or graduality. Many recent works have tackled verifying
  these properties, but the resulting mathematical developments are
  highly repetitive and tedious, with few reusable theorems persisting
  across different developments.

@@

\newif\ifdraft
\drafttrue
\renewcommand{\max}[1]{\ifdraft{\color{blue}[{\bf Max}: #1]}\fi}
\newcommand{\eric}[1]{\ifdraft{\color{orange}[{\bf Eric}: #1]}\fi}
\newcommand{\tingting}[1]{\ifdraft{\color{red}[{\bf Tingting}: #1]}\fi}

@@

% \input{syntax}

% \input{semantics}

\input{terms}

\input{ordering}

@@

\input{categorical-model}

%\input{denotational-model}
\input{concrete-model}

\input{discussion}

\appendix
\input{appendix}

% \section{Discussion}\label{sec:discussion}
